---
title: "Data Manipulation"
author: "Claire Liu"
output: github_document
---

This file is for doing data import.

```{r}
library(tidyverse)
library(readxl)
library(haven)
```

Import our first dataset.

```{r}
litters_df = 
  read_csv(file = "data/FAS_litters.csv",
           na = c("","NA", ".")
           )

litters_df = janitor::clean_names(litters_df)

pups_df = 
  read_csv(
    file = "data/FAS_pups.csv", 
    skip = 3,
    na = c("", "NA", ".")
  )

pups_df = janitor::clean_names(pups_df)
```


```{r}
select(litters_df, group, litter_number, gd0_weight, pups_born_alive)

select(litters_df, group:gd_of_birth)

select(litters_df, -pups_survive)

select(litters_df, starts_with("gd"), group)

select(litters_df, LITTER = litter_number, everything())

select(litters_df, starts_with("gd"))
```

a coupe of alternatives

```{r}
rename(litters_df, GROUP = group)

relocate(litters_df, litter_number, gd0_weight)
```

## `filter` some rows

```{r}
filter(litters_df, gd_of_birth == 19)

filter(litters_df, pups_born_alive <= 5)

filter(litters_df, pups_born_alive != 6)

filter(litters_df, group %in% c("Con7", "Con8"))

filter(litters_df, group %in% c("Con7", "Con8"), pups_born_alive < 7)

filter(litters_df, pups_born_alive < 4 | pups_born_alive > 8)

drop_na(litters_df, gd0_weight)
```

In the pups dataset, keep only rows where PD walk is less than 11 and sex is 2

```{r}
filter(pups_df, pd_walk < 11, sex == 2)
```


## `mutate` to change some columns

```{r}
mutate(
  litters_df,
  wt_gain = gd18_weight - gd0_weight,
  wt_gain_squared = wt_gain ^ 2,
  group = str_to_lower(group)
)
```

In the pups datasset, create a variabele that subtracts 7 from PD walk

```{r}
mutate(pups_df, df_walk_minus_7 = pd_walk - 7)
```

## `arrange` the data

```{r}
arrange(litters_df, pups_born_alive)
arrange(litters_df, desc(pups_born_alive))
```

## Do manipalation steps...

This is bad

```{r}
# litters_df = read_csv("data/FAS_litter.csv", na = c("", "NA", "."))
# litters_rename_df = janitor::clean_names(litter_df)
# litters_with_gd_df = select(litters_rename_df, group, starts_with("gd"))
# litters_no_na_df = drop_na(litters_with_gd_df)
# litters_wt_gain_df = mutate(litters_no_na_df, wt_gain = gd18_weight - gd0_weight)
```

This is worse
...

This is good

```{r}
litters_df =
  read_csv("data/FAS_litters.csv", na = c("", "NA", ".")) |>
  janitor::clean_names() |> 
  select(group, starts_with("gd")) |> 
  drop_na() |> 
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group)
  )
```

load pups, clean names, drop missing, keep group and pd varaibles, add pd walk -7

